---
title: "6372 Auto Project"
author: "Justin Ehly"
date: "1/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
### Libraries ###
library(car)
library(magrittr)
library(tibble)
library(naniar)
library(VIM)
library(FactoMineR)
library(purrr)
library(gdata)
library(glmnet)
library(leaps)
library(ISLR)
library(caret)
library(tidyverse)
library(stringr)
library(ggplot2)
library(GGally)
library(scales)
library(dplyr)
```

```{r get-data}
### Download Data and Clean ###

# teammates that fork this code chunk, be sure to set
# your working directory to the DataSet folder in your your github folder
#setwd("C:/Users/justi/Documents/GitHub/6372-Auto-Pricing-Project/DataSet")
setwd(choose.dir()) #this only works on a windows machine

auto_data <- read.csv("data1.csv")
#view(auto_data)

summary(auto_data)
str(auto_data)
names(auto_data)
autos <- auto_data
```

```{r clean data}
#############################

###    Clean the dataset  ###

#############################

autos <- rename(autos, mpg.highway = highway.MPG, mpg.city = city.mpg)
# replace all chr with factors
autos[sapply(autos,is.character)] <- lapply(autos[sapply(autos,is.character)], as.factor)
autos$Year <- as.factor(autos$Year) # make Year into a factor
autos$Number.of.Doors <- as.factor(autos$Number.of.Doors) # change to factor

str(autos)
#view(autos)


### Audi A6 sure does have an awesome mpg.highway - should be 34 not 354 ###
autos$mpg.highway[1120] <- 34

# create an average mpg column
autos$avgMPG <- (autos$mpg.city+autos$mpg.highway) / 2
names(autos)
autos <- autos[,-c(13,14)]
str(autos)

##### Look for any missing values including NA, variants of NA and 'UNKNOWN's in the data set ####

missing <- sapply(autos, function(x) sum(sum(x %in% common_na_strings), 
                                        sum(x %in% common_na_numbers),
                                        sum(is.na(x)), 
                                        sum(x %in% 'UNKNOWN')))
missing 

### Plot missing data ### 
missing <- data.frame(as.list(missing)) # convert names int vector to dataframe
missing <- as.data.frame(t(as.matrix(missing))) # transpose the matrix
missing <- rename(missing, 'MissingValues'='V1')
missing$Variables <- rownames(missing)
missing <- missing[,c(2,1)]
rownames(missing) <- 1:nrow(missing)
missing <- missing[order(missing$MissingValues, decreasing = TRUE),]
missing

missing %>% ggplot(aes(y=reorder(Variables,MissingValues), 
                    x=MissingValues, fill=Variables)) + 
  geom_col(show.legend = FALSE) + 
  labs (title ="Missing Values by Variable",
        x = "No. Missing Values",
        y = "Varaiables") +
  geom_text(aes(label=comma(MissingValues, accuracy = 1)), nudge_x = 100, size=3) +
  scale_x_continuous(labels=comma) +
  theme_bw()

### Clean up data ###

#Fiat 500e - electric so no cylinders
autos$Engine.HP[c(540:542)] <- 111 

#2017 Continental
autos[c(2906:2909),]
autos$Engine.HP[c(2906:2909)] <- 400 

#2017 Escape
autos[c(4204:4207),]
autos$Engine.HP[c(4204:4207)] <- 179 

#2013-2014 Fit EV
autos[c(4706,4707),]
autos$Engine.HP[c(4706,4707)] <- 123

#2015 Ford Focus EV 143hp
autos[c(4786,4790,4799),]
autos$Engine.HP[c(4786,4790,4799)] <- 143

#2005 Ford Freestar only vans above $29k have 201 HP
autos[c(4915:4928),] %>% select(Engine.HP,avgMPG, MSRP) %>% arrange(avgMPG)
autos$Engine.HP[c(4915:4918)] <- 193
autos$Engine.HP[c(4919:4920)] <- 201

# 2014 Mitsubishi i-MiEV
autos$Engine.HP[5779] <- 66

# 2015-2016 Kia Soul EV
autos$Engine.HP[c(9851:9855)] <- 109

#2013-2014 Toyota Rav4 EV
autos$Engine.HP[c(8375:8376)] <- 154

#Telsa Model S missing values
autos %>% filter(Make == "Tesla")

tesla <- read.csv("tesla.csv")
#view(tesla)

# mass replace tesla missing values since they were mostly all independent
for(i in 6922:6939){autos$Engine.HP[i] <- tesla$Engine.HP[i-(6921)]}
for(i in 6922:6939){autos$Number.of.Doors[i] <- tesla$Number.of.Doors[i-(6921)]}
#view(autos)

# 2017 Lincoln MKZ - all FWD have 240hp
autos$Engine.HP[c(6909,6911,6917,6919)] <- 240

# 2015 Mercedes M-Class Diesel
autos$Engine.HP[6579] <- 200

#2014-2016 Nissan Leaf - all 107 hp
autos$Engine.HP[c(6386:6395)] <- 107

#### Work on missing cylinders 
missingCyl <- which (is.na(autos$Engine.Cylinders))
#write.csv(autos[missingCyl,],"MissingCyl.csv")
#view(autos[missingCyl,])
#change electric cars to 'e' for cylinders since they don't have any
autos$Engine.Cylinders <- ifelse(autos$Engine.Fuel.Type == 'electric','E',autos$Engine.Cylinders)
#change the mazda RX cars to 'R' for rotary engine since they don't have cylinders
autos$Engine.Cylinders[c(8696:8715)] <- 'R'

# any remaining missing values?
sapply(autos, function(x) sum(sum(x %in% common_na_strings), 
                                        sum(x %in% common_na_numbers),
                                        sum(is.na(x)), 
                                        sum(x %in% 'UNKNOWN')))

# the software seems to think there is 1 value missing for number.of.doors
autos[which(is.na(autos$Number.of.Doors)),]
# 2- door ferrari ff
autos$Number.of.Doors[which(is.na(autos$Number.of.Doors))] <-2

# Engine.Fuel.Type - suzuki is missing
autos$Engine.Fuel.Type[c(11322:11324)] <- 'regular unleaded'

# All that is left now is to either predict or impute the Market.Category - 
# rather, let's follow the path of the directions and just create and "exotic" attribute
exotic <- c('Ferrari','Alfa Romeo','McLaren', 'Maybach', 'Porsche', 
            'Bentley', 'Lamborghini', 'Spyker', 'Rolls-Royce', 'Maserati',
            'Aston Martin', 'Lotus', 'Bugatti')
for(i in 1:length(autos$Make)){
  ifelse(autos$Make[i] %in% exotic,autos$Exotic[i] <- 'Exotic', autos$Exotic[i] <- 'Not Exotic')
}
# Actually after some further evaluation, I feel setting the exotic car price ta at $100k is a better fit for analysis also, let's remove the Bugatti, those literally sale <100 annually

autos<-autos[-c(11363:11365),]
unique(autos$Make)

autos$Exotic <- as.character(autos$Exotic)
autos$Exotic <- case_when(
  autos$MSRP > 100000 ~ "Exotic",
  TRUE ~ autos$Exotic
)

#view(autos %>% filter(autos$MSRP > 100000) )

autos$Exotic <- as.factor(autos$Exotic)
str(autos)
autos$Market.Category <- NULL #remove this unused column, it's overly complex

autos$Engine.HP <- as.integer(autos$Engine.HP)

#2015 Impala duel-fuel - since only a 30hp difference, might as well use the higher hp

autos[c(5826,5831, 5832, 5834, 5840, 5841),]
autos$Engine.HP[c(5826,5831,5832,5834,5840, 5841)] <- 260

str(autos)
# replace all chr with factors
autos[sapply(autos,is.character)] <- lapply(autos[sapply(autos,is.character)], as.factor)
autos$Year <- as.factor(autos$Year) # make Year into a factor
autos$Number.of.Doors <- as.factor(autos$Number.of.Doors) # change to factor
autos[sapply(autos,is.integer)] <- lapply(autos[sapply(autos,is.integer)], as.numeric)


# Transmission Types has 19 missing values
autos$Transmission.Type[c(1290:1291)] <- "AUTOMATIC" #Oldsmobile Achieva
autos$Transmission.Type[c(4692:4694)] <- "AUTOMATIC" #Pontiac Firebird with 15/23 MPG and V8
autos$Transmission.Type[c(6159,6161,6166,6175)] <- "AUTOMATIC" #1999/2000 GMC Jimmy
autos$Transmission.Type[c(6367,6369)] <- "AUTOMATIC" #1993 Chrysler LeBaron
autos$Transmission.Type[c(8043,8044,8047:8050,8052,8054)] <- "AUTOMATIC" #1991 Dodge Ram 150

# recheck for missing values
sapply(autos, function(x) sum(sum(x %in% common_na_strings), 
                                        sum(x %in% common_na_numbers),
                                        sum(is.na(x)), 
                                        sum(x %in% 'UNKNOWN')))
# Note: verified that there are no missing values in Engine.HP


# export the final cleansed data set for reporting
# write.csv(capture.output(str(autos)), "cleanData.csv")

auto_data_clean <- autos  # easy starting over point with the clean data set


```



```{r objective 1 EDA}
####################

### EDA  ###########

####################
autos <- auto_data_clean # starting fresh with clean data in case you need to rerun this again

# Interested to see how the MSRP evolves over time
autos %>% select(MSRP, Year) %>% filter(MSRP < 100000) %>%
  ggplot(aes(x=Year, y=MSRP, fill=Year)) +
  geom_boxplot(aes(group = Year), show.legend = FALSE) +
  labs(title = "MSRP Ranges by Year",
       y = "MSRP", 
       x = "Model Year") +
  scale_y_continuous(label=comma) +
  theme_bw()

# based on this graph, we can separate cars pre 2001 from cars post 2001,
# creating a new variable to address this

autos$Year <- as.numeric(as.character(autos$Year))
autos$Year <- case_when( 
    autos$Year < 2001 ~ 2000,
    TRUE ~ autos$Year)

autos <- autos[,c(13,1:12,14:15)] #put MSRP first

autos %>% select(Year, MSRP) %>% group_by(Year) %>%
  summarize(max=max(MSRP),mean=mean(MSRP), min=min(MSRP))

# seems logical to remove anything before 2001 because the values don't make logical sense...did car prices suddenly make a jump from a mean of $2,530 to $41,501?
# dropping those older cars

autos <- autos %>% filter(Year > 2000)

autos$Year <- as.factor(autos$Year)

# Interested to see how the MSRP look without models older than 2001
autos %>% select(MSRP, Year) %>% filter(MSRP < 100000) %>%
  ggplot(aes(x=Year, y=MSRP, fill=Year)) +
  geom_boxplot(aes(group = Year), show.legend = FALSE) +
  labs(title = "MSRP Ranges by Year with everyone before 2001 grouped",
       y = "MSRP", 
       x = "Model Year") +
  scale_y_continuous(label=comma) +
  theme_bw()

summary(autos) #summary of main autos df


### What's going on with the Popularity variable? ###

### Make ###
t(aggregate(Popularity~Make,data=auto_data,min)) #appears each make has a specific popularity
autos %>% dplyr::group_by(Make) %>% 
  summarise(mean = mean(Popularity), range=range(Popularity)) %>%
  ggplot(aes(x=Make, y=mean, fill=Make)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Average Popularity Score by Make",
       y = "Averge Popularity", 
       x = "Car Make") + 
  scale_y_continuous(label=comma) +
  coord_flip() +
  theme_bw()

#Seems each make just has one popularity score
auto_data %>% dplyr::group_by(Make) %>% 
       summarize(min=min(Popularity),
                 mean=mean(Popularity), 
                 median=median(Popularity),
                 max=max(Popularity))
# Confirmed, Popularity is directly tied to Make, they are interchangeable and thus we only need to use one of them

# remove popularity
autos <- autos[,-13]
names(autos)


##### Relationship between Make and MSRP ###

autos %>% filter(Exotic == "Not Exotic") %>%
  ggplot(aes(x=reorder(Make, desc(Make)), y=MSRP, fill=Make)) +
  geom_boxplot(aes(group = Make), show.legend = FALSE) +
  labs(title = "Non-Exotic Make by MSRP",
       y = "MSRP", 
       x = "Car Make") +
  scale_y_continuous(label=dollar_format()) +
  coord_flip() +
  theme_bw()

autos %>% filter(Exotic == "Exotic") %>%
  ggplot(aes(x=reorder(Make, desc(Make)), y=MSRP, fill=Make)) +
  geom_boxplot(aes(group = Make), show.legend = FALSE) +
  labs(title = "Exotic Make by MSRP",
       y = "MSRP", 
       x = "Car Make") +
  scale_y_continuous(label=dollar_format()) +
  coord_flip() +
  theme_bw()

### Relationship between Cylinders and Horsepower ##
autos %>% #filter(Exotic == "Not Exotic") %>%
  ggplot(aes(x=reorder(Engine.Cylinders, Engine.HP), 
             y=Engine.HP, fill=Engine.Cylinders)) +
  geom_boxplot(aes(group = Engine.Cylinders), show.legend = FALSE) +
  labs(title = "Engine Cylinders vs Horsepower",
       y = "Horsepower", 
       x = "Cylinders/ Engine Type",
       caption = "R: RotaryEengine; E: Electric Engine") +
  scale_y_continuous(label=comma) +
  coord_flip() +
  theme_bw()

# look at the matrix plots of explanatory variables compared to MSRP
# keep in mind we are reduced to just 2 - Engine.HP and avgMPG, but adding in Year.Split
# to show the difference when we separate the between before and after 2000/2001 cars
autos %>% select(MSRP, Engine.HP, avgMPG) %>% 
  filter(MSRP < 1000000) %>% ggpairs() + 
  labs(title = "GGPairs Chart - MSRP > $1,000,000 withheld") + theme_bw()


autos_EDA_stop <- autos # setting another quick restart point

```


```{r clean up the categorical variables}
autos <- autos_EDA_stop

aov.mod <- aov(MSRP~.-Model, data=autos)
summary(aov.mod)

# reran AOV, all are significant

names(autos)
autos <- autos[,-c(10,13)] #removes doors, avgMPG

# to simpilify - going to remove models for the lm
pre.mod <- lm(MSRP~.-Model, data = autos)
summary(pre.mod)
# reference = Acura
# Let's mix up the makes based on the insignif p-value >0.05

autos$Make <- as.character(autos$Make)

autos$Make <- case_when(
  autos$Make == "Acura" ~ "Lux",
  autos$Make == "Alfa Romeo" ~ "Euro",
  autos$Make == "Audi" ~ "Euro_Lux",
  autos$Make == "BMW" ~ "Euro_Lux",
  autos$Make == "Cadillac" ~ "Lux",
  autos$Make == "FIAT" ~ "Euro",
  autos$Make == "Genesis" ~ "Lux",
  autos$Make == "HUMMER" ~ "Lux",
  autos$Make == "Infiniti" ~ "Lux",
  autos$Make == "Lexus" ~ "Lux",
  autos$Make == "Lincoln" ~ "Lux",
  autos$Make == "Mercedes-Benz" ~ "Euro_Lux",
  autos$Make == "Saab" ~ "Lux",
  TRUE ~ autos$Make)

autos$Make <- as.factor(autos$Make)

pre.mod <- lm(MSRP~.-Model, data = autos)
summary(pre.mod)
# Plymouth is still standing out, but barely - we will leave it alone because i dont have anything to bundle it with 




##################

### Fuel Types ###

##################
# test for differences in fuel types related to MSRP
test_lm <- lm(MSRP~. - Model, data=autos)
summary(test_lm)


# Testing for unleaded differences
test_lm <- autos %>% filter(grepl("unleaded",Engine.Fuel.Type)) %>%
  lm(MSRP~Engine.Fuel.Type, data=.)
summary(test_lm)


autos$new.fuel <- as.factor(autos$Engine.Fuel.Type)
autos$new.fuel <- as.factor(autos$new.fuel)

autos$new.fuel <- str_replace_all(autos$new.fuel, "premium unleaded \\(recommended\\)", "regular unleaded")
unique(autos$new.fuel)
test_lm <- lm(MSRP~new.fuel, data=autos)
summary(test_lm)

autos$new.fuel <- str_replace_all(autos$new.fuel, "flex-fuel \\(premium unleaded recommended\\/E85\\)", "flex-fuel (unleaded/E85)")
test_lm <- lm(MSRP~new.fuel, data=autos)
summary(test_lm)

autos$new.fuel <- str_replace_all(autos$new.fuel, "flex-fuel \\(unleaded\\/E85\\)", "flex-fuel")
autos$new.fuel <- str_replace_all(autos$new.fuel, "flex-fuel \\(unleaded\\/natural gas\\)", "flex-fuel")
test_lm <- lm(MSRP~new.fuel, data=autos)
summary(test_lm)

autos$new.fuel <- case_when(
  autos$new.fuel == "natural gas" ~ "alternative",
  autos$new.fuel == "flex-fuel" ~ "alternative",
  autos$new.fuel == "diesel" ~ "alternative",
  autos$new.fuel == "electric" ~ "alternative",
  TRUE~autos$new.fuel)

autos$new.fuel <- as.factor(autos$new.fuel)
levels(autos$new.fuel)

autos$Engine.Fuel.Type <- autos$new.fuel
levels(autos$Engine.Fuel.Type)
str(autos)
names(autos)
new.fuel.index <- grep("new.fuel", colnames(autos))
# remove new.fuel
autos <- autos[,-new.fuel.index]


###### New LM test ##########

test_lm <- lm(MSRP~.-Model, data=autos)
summary(test_lm)
names(autos)

# Engine.HP looks good pvalue = < 2e-16
# Transmission needs some cleaning up

#########################

### Transmission  #######

#########################

# Transmission Direct Drive appears insig pvalue =0.6314
levels(autos$Transmission.Type)

autos$Transmission.Type <- str_replace_all(autos$Transmission.Type, "AUTOMATED_MANUAL", "n")
autos$Transmission.Type <- str_replace_all(autos$Transmission.Type, "DIRECT_DRIVE", "n")
autos$Transmission.Type <- str_replace_all(autos$Transmission.Type, "n", "AUTO-MANUAL-OR-DIRECT-DRIVE")

autos$Transmission.Type <- as.factor(autos$Transmission.Type)

### new LM test ###

test_lm <- lm(MSRP~.-Model, data=autos)
summary(test_lm)

# Transmisson.Type looks good now
# Wheels Driven needs some attention

############################

### Driven Wheels ##########

############################

# Driven Wheels
autos$Driven_Wheels <- str_replace_all(autos$Driven_Wheels, "four wheel drive", "zzz")
autos$Driven_Wheels <- str_replace_all(autos$Driven_Wheels, "all wheel drive", "zzz")
autos$Driven_Wheels <- str_replace_all(autos$Driven_Wheels, "front wheel drive", "zzz")
autos$Driven_Wheels <- str_replace_all(autos$Driven_Wheels, "zzz", "all or 4 wheel drive")
autos$Driven_Wheels <- as.factor(autos$Driven_Wheels)
levels(autos$Driven_Wheels)

test_lm <- lm(MSRP~. - Model, data=autos)
summary(test_lm)

# Vehicle size looks ok too
# Vehicle style may need some attention


# convertible suv's seem to be insignificant, let's just check to see what we can marry them to

autos %>% filter(Vehicle.Style == "Convertible SUV") %>% distinct(Make, Model, Year) %>% arrange(Make)
# chevy, landrovers, nissans, suzukis are 2-door

autos$Vehicle.Style <- as.character(autos$Vehicle.Style)
autos$Vehicle.Style <- case_when(
  autos$Vehicle.Style == "2dr SUV" ~ "SUV",
  autos$Vehicle.Style == "4dr SUV" ~ "SUV",
  autos$Vehicle.Style == "Convertible SUV" ~ "SUV",
  autos$Vehicle.Style == "4dr Hatchback" ~ "Sedan",
  autos$Vehicle.Style == "2dr Hatchback" ~ "Sedan",
  autos$Vehicle.Style == "Wagon" ~ "Sedan",
  autos$Vehicle.Style == "Extended Cab Pickup" ~ "Pickup",
  autos$Vehicle.Style == "Regular Cab Pickup" ~ "Pickup",
  autos$Vehicle.Style == "Crew Cab Pickup" ~ "Pickup",
  autos$Vehicle.Style == "Cargo Minivan" ~ "Cargo",
  autos$Vehicle.Style == "Cargo Van" ~ "Cargo",
  autos$Vehicle.Style == "Passenger Van" ~ "Passenger Van",
  autos$Vehicle.Style == "Passenger Minivan" ~ "Passenger Van",
  TRUE ~ autos$Vehicle.Style
)

autos$Vehicle.Style <- as.factor(autos$Vehicle.Style)
levels(autos$Vehicle.Style)

### Lm test
test_lm <- lm(MSRP~., data=autos)
summary(test_lm)

# looks good now


vif(test_lm) # all vif's look good around 1

vif(test_lm)[,3]^2
## (avgMPG may be an issue, but let's hang tight with that)

autos$Make <- autos$Make
autos$Model <- autos$Model
autos$Year <- autos$Year
autos <- autos
str(autos)

str(autos)
summary(autos)

mod <- lm(MSRP~ .-Make-Model,autos)
summary(mod)

```




```



```{r create 3 data sets}
##############################

#### Create 3 Data Sets  #####

##############################

set.seed(123)
spec = c(train = .8, test = .1, validate = .1) #set the split percentages 80/10/10

ind = sample(cut(
  seq(nrow(autos)), 
  nrow(autos)*cumsum(c(0,spec)),
  labels = names(spec)
))

autosplits = split(autos, ind)

# check results
sapply(autosplits, nrow)/nrow(autos)
#     train       test   validate 
# 0.79998321 0.09996643 0.10005036 

# assign simple common names to each dataset 
train <- autosplits$train
test <- autosplits$test
validate <- autosplits$validate

str(train)
str(test)
str(validate)


```

Objective 1: Display the ability to build regression models using the skills and discussions from Unit 1 and 2 with the purpose of identifying key relationships and interpreting those relationships. A key question of interest that must be addressed in this analysis is the importance of the "Popularity" variable. While the details of this variable are vague, it was created from social media, and the "higher ups" are curious how much general popularity can play a role in the retail price of a vehicle.

Since we simply have too many predictors to run a forward, backward or stepwise test
2^11 = 2,048 possible models

Best to jump into LASSO for variable selection


### ###################

### LASSO MODEL ###

### ###################

```{r LASSO model}

# formatting data for GLM net
x = model.matrix(MSRP~. -Make -Model,train)[,-1]
y = train$MSRP

xtest <- model.matrix(MSRP~. -Make -Model,test)[,-1]
ytest <- test$MSRP

grid <- 10^seq(10,-2, length = 100)
lasso.mod <- glmnet(x, y, alpha=1, lambda = grid)

cv.out <- cv.glmnet(x, y, alpha=1) # alpha = 1 performs LASSO

bestlambda <- cv.out$lambda.min #Optimal penalty parameter. We can also make this call visually

lasso.pred <- predict(lasso.mod, s= bestlambda, newx=xtest)

testMSE_LASSO <- mean((ytest-lasso.pred)^2)
testMSE_LASSO

coef(lasso.mod, s=bestlambda)

### Plot LASSO results ###
par(mfrow=c(2,2))
plot(lasso.mod)
plot(lasso.mod,
     xvar = "lambda")
plot(cv.out)

```
